\subsubsection{Single-Point Crossover}
\label{sec:keen:op:cx:single_point}

  Single-point crossover is one of the most foundational and widely utilized 
  crossover methods in the domain of EC. Rooted in its simplicity and 
  efficiency, this method selects a random \enquote{crossover point} on the 
  input chromosomes and exchanges the subsequences of genes situated after this 
  point between the two input chromosomes, producing two outputs.

  \begin{definition}[Single-Point Crossover]
    The \textit{single-point crossover} operator takes a pair of input 
    chromosomes and randomly selects a crossover point on them. It then 
    exchanges the subsequences of genes situated after this point between the 
    two input chromosomes to produce two outputs. More formally, a single-point 
    crossover is represented as:

    \begin{equation}
        X_\mathrm{sp} :\: 
            \mathbb{P} \times [0,\, 1] \times [0,\, 1] \times \{0,\, 1\} \rightarrow \mathbb{P};\;
            (P,\, \rho_i,\, \rho_\mathbf{c}, e) \mapsto X_\mathrm{sp}(P,\, \rho_i,\, \rho_\mathbf{c}, e)
    \end{equation}

    where:

    \begin{itemize}
        \item \(P\) represents a population of binary chromosomes.
        \item \(\rho_i\) symbolizes the likelihood of applying the crossover to 
            an individual.
        \item \(\rho_\mathbf{c}\) symbolizes the likelihood of applying the 
            crossover to a chromosome.
        \item \(e\) is a boolean value that determines whether the same individual can be used as both parents.
    \end{itemize}
  \end{definition}

  \begin{remark}
    The simplicity of single-point crossover makes it a popular choice in many 
    EC applications. However, it might not always ensure adequate exploration 
    of the search space, especially in problems where gene positions have 
    strong interactions. In these scenarios, permutation crossover methods like 
    \textit{ordered crossover} and \textit{partially mapped crossover} might be 
    more effective.
  \end{remark}

  For better clarity on the single-point crossover operation, see the graphical 
  representation in \vref{fig:bg:ga:var:cx:single_point}, which showcases the 
  random selection of a crossover point and the subsequent exchange of gene 
  subsequences.

  Note: There exists a generalized version of single-point crossover named 
  \textit{multi-point crossover}. This method selects multiple crossover points 
  on the input chromosomes and exchanges the subsequences of genes situated 
  after these points between the two input chromosomes to produce two 
  outputs. However, this method is currently not supported in \textit{Keen}.
