\section{Introduction}
\label{sec:introduction}
  Within the expansive landscape of \textit{evolutionary computation} (EC), a
  multitude of methodologies and algorithms have emerged, each demonstrating
  significant success in addressing complex optimization problems.
  Nevertheless, the practical application of these techniques often presents
  its own challenges, most notably, the extensive repetitive coding required
  and the absence of an intuitive, user-friendly platform.
  To address these challenges, this chapter introduces \emph{Keen}, an
  innovative genetic algorithms framework developed using the \textit{Kotlin}
  programming language.

  As a comprehensive and adaptable framework, Keen embarks on an ambitious
  journey to promote the application and acceptance of evolutionary algorithms
  across various domains.
  This well-structured framework is currently designed to simplify the
  implementation process of \textit{genetic algorithms} (GAs) and
  \textit{genetic programming} (GP), with an emphasis on allowing extension to
  other evolutionary algorithms, thereby accelerating the pace of research
  and development in the EC field.

  This chapter provides an in-depth examination of the \textit{Keen
  framework}, illuminating its architecture, fundamental components, and the
  array of functionalities it offers.
  We explore the range of genetic operators within Keen, including various
  selection, mutation, and crossover techniques.
  A comprehensive account of the GAs and GP methods within \textit{Keen}'s
  repertoire will be provided, underscoring their distinctive attributes and
  potential use-cases.
  In addition, the chapter delves into the framework's parallelism
  capabilities, a feature that significantly boosts computational efficiency
  and scalability.

  Beyond its existing features, the extensibility of the Keen framework will
  also be examined, demonstrating its readiness to incorporate novel
  algorithms and techniques.
  Keen is designed with a view to facilitate easy expansion and modification
  of its components, encouraging innovation and continuous progression, in
  tandem with advancements in the EC field.

  The introduction of \textit{Keen} marks a valuable contribution to both the
  \textit{Kotlin} and AI communities, offering a robust, user-friendly
  platform for the development, testing, and deployment of evolutionary
  algorithms.
  We envision that this pioneering framework will encourage broader
  application of evolutionary algorithms, thus pushing the boundaries of
  artificial intelligence research and application.