\chapter{Glossary}
\label{appendix:glossary}
  In this appendix we present a glossary of terms used throughout this document 
  that may be unfamiliar to the reader and are not defined in the main text.

  \subimport{./}{A.tex}

  \section*{C}
    \begin{definition}[Chromosome]
    \label{def:chromosome}
      Representation of a single column of genetic information of a candidate solution to a given
      optimization problem.

      Formally, a chromosome is a vector \(\textbf{c} = (g_1, g_2, \dots, g_n)\), where \(g_i\) is a
      gene (see \vref{def:gene}).
    \end{definition}

  \section*{E}
    \begin{definition}[Ephemeral Constant]
    \label{def:ephemeral_constant}
      A constant that is randomly generated at the start of the program and 
      remains constant throughout the execution of the program.

      This is used to represent constant values in the program.
    \end{definition}

    \begin{definition}[Evolutionary computation]
    \label{def:evolutionary_computation}
      Family of algorithms for global optimization inspired by the process of natural selection.

      This typically involves processes mimicking natural selection, mutation, recombination, and 
      survival of the fittest.

      The solutions to a problem are encoded as a set of \enquote{individuals} in a 
      \enquote{population}.

      Over multiple generations, these individuals are selected and modified (via genetic operators 
      like crossover and mutation) in order to find better solutions.
    \end{definition}
    
  \section*{G}
    \begin{definition}[Gene]
    \label{def:gene}
      Representation of a single component of a candidate solution to a given optimization problem.

      Formally, for a multi-dimensional function \(f\), a gene is an element \(g\) in the domain of
      \(f\).
    \end{definition}

    \begin{definition}[Generation]
    \label{def:generation}
      Number of iterations performed by an evolutionary algorithm.
    \end{definition}

    \begin{definition}[Genotype]
    \label{def:genotype}
      Representation of the full genetic information of a candidate solution to a given optimization
      problem.

      Formally, a genotype is a matrix \(\mathbf{G} = (\textbf{c}_1, \textbf{c}_2, \dots, 
      \textbf{c}_n)\), where \(\textbf{c}_i\) is a chromosome (see \vref{def:chromosome}).
    \end{definition}

  \section*{I}
    \begin{definition}[Individual]
    \label{def:individual}
      A candidate solution to a given optimization problem.

      Formally, an individual is a pair \((\mathbf{G}, \mathbf{f})\), where \(\mathbf{G}\) is the
      genotype (see \vref{def:genotype}) and \(\mathbf{f}\) is the fitness value of the
      individual.
    \end{definition}

  \subimport{./}{M.tex}

  \section*{P}
    \begin{definition}[Parameter optimization]
    \label{def:parameter_optimization}
      Optimization problem where the solution is a set of parameters that 
      optimize a given function.
    \end{definition}

    \begin{definition}[Phenotype]
    \label{def:phenotype}
      Same as \vref{def:individual}.
    \end{definition}

    \begin{definition}[Population]
    \label{def:population}
      Set of candidate solutions to a given optimization problem.
    \end{definition}

    \begin{definition}[Program induction]
    \label{def:program_induction}
      Inference of an algorithm or program featuring recursive calls or 
      repetition control structures, starting from information that is known to 
      be incomplete, called the evidence, such as positive and negative I/O 
      examples or clausal constraints.
    \end{definition}

  \section*{S}
    \begin{definition}[Search space]
    \label{def:search_space}
      Set of all candidate solutions to a given optimization problem.
    \end{definition}

    \begin{definition}[Selector]
      See \vref{def:selection_operator}.
    \end{definition}

  \section*{U}
    \begin{definition}[Unimodal function]
    \label{def:unimodal_function}
      A unimodal function is a function that, within a given domain or interval,
      has only one local maximum and one local minimum, which can also be the
      global maximum and minimum respectively.
      In other words, the function increases to a certain point and then 
      decreases, or vice versa.
      This single peak or trough is the \enquote{uni} in \enquote{unimodal}.

      Mathematically speaking, if \(f\) is a unimodal function in the interval 
      \([a,\, b]\), then there exists some point \(c\) in \([a,\, b]\) such 
      that:

      \begin{enumerate}
        \item \(f(x)\) is increasing on \([a,\, c]\) and decreasing on \([c,\, 
          b]\), or
        \item \(f(x)\) is decreasing on \([a,\, c]\) and increasing on \([c,\,
          b]\).
      \end{enumerate}
    \end{definition}

  \section*{V}
    \begin{definition}[Variadic function]
    \label{def:variadic_function}
      Function that accepts a variable number of arguments.
    \end{definition}
