% \setabstractname { Summary }
\begin{abstract}
    This thesis explores Evolutionary Computation (EC) within Artificial Intelligence, particularly focusing on its 
    application in software engineering and scientific computing. It introduces an EC framework, highlighting its 
    modularity and ability to integrate various EC algorithms. The framework's efficacy is demonstrated through case 
    studies on Genetic Algorithms and Linear Genetic Programming, emphasizing its ability to solve complex 
    computational problems.

    The development is grounded in a detailed review of existing EC frameworks, with an emphasis on their 
    functionalities and architectural details. Additionally, the thesis explores the theoretical dimensions of EC, 
    providing a comprehensive analysis of its fundamental concepts and methodologies. This study enriches the 
    understanding of EC and contributes to academic discussions, establishing a firm foundation for future research and 
    applications.

    The central hypothesis suggests that a framework emphasizing modularity and extensibility can positively impact 
    evolutionary computation. This framework is expected to be user-friendly and flexible, potentially streamlining 
    evolutionary algorithm implementations and fostering efficient research and innovation.

    The structure of the thesis is methodically laid out as follows:
    \begin{enumerate}
        \item \textbf{Background}: Establishes the theoretical foundation for the central concepts.
        \item \textbf{State of the Art}: Reviews current advancements and frameworks in EC.
        \item \textbf{Framework Design and Implementation}: Details the architecture and implementation of the 
            framework.
        \item \textbf{Function Optimization Case Study}: Tests the framework with real-world function optimization
            problems.
        \item \textbf{Crash Reproduction Problem Case Study}: Demonstrates the framework's application in another 
            domain.
        \item \textbf{Conclusion and Future Work}: Summarizes the work and proposes future research directions.
    \end{enumerate}

    This structure ensures a comprehensive understanding of the research's objectives, methodology, and outcomes, providing a coherent narrative throughout the thesis.
\end{abstract}