\chapter{State of the Art: Relevant Work}
\label{chap:sota}
  This chapter delves into the current state of the art in evolutionary 
  computation frameworks.
  While evolutionary algorithms themselves aren't new, their applications keep 
  evolving\footnote{No pun intended.} with technological advancements in 
  programming languages and tools.

  We'll spotlight five major frameworks: \emph{Agile Artificial Intelligence in 
  Pharo}, \emph{DEAP}, \emph{Jenetics}, \emph{ECJ}, and \emph{GeneticSharp}.
  Each section provides an overview of the respective framework, complete with 
  code samples, showcasing their syntax and distinct features.
  This exploration aims to give readers a broad understanding of the array of 
  tools in this domain.

  Moreover, this review sets the context for our forthcoming contribution: an 
  evolutionary computation framework crafted in Kotlin.
  To relate, \vref{lst:app:keen_omp} features the \emph{Keen} framework's 
  approach to the \emph{OneMax} problem, akin to the examples herein.

  While we strive for objectivity, we recognize that personal biases may color 
  our perceptions.
  We don't position ourselves as experts in the covered frameworks or languages, 
  and there's a possibility that the provided implementations could be optimized 
  further.
  We invite readers to delve into these frameworks firsthand and shape their own 
  conclusions.

  For transparency, it's worth noting that this review is impartial.
  We haven't received any compensation from, nor are we affiliated with, any of 
  the framework developers aside from Bergel's \emph{Agile Artificial 
  Intelligence in Pharo} book.
  This chapter reflects our individual insights and experiences.

  \subimport{.}{OMP.tex}
  \subimport{./}{Pharo.tex}
  \subimport{.}{DEAP.tex}
  \subimport{.}{Jenetics.tex}
  \subimport{.}{ECJ.tex}
  \subimport{.}{GeneticSharp.tex}
  \subimport{.}{Other.tex}
  \subimport{.}{Limitations.tex}
