\section{Hypothesis and Research Questions}
\label{sec:hyphotesis_and_research_questions}
  \subsection{Hypothesis}
    We propose that developing an evolutionary algorithms framework focused on extensibility and modularity, could 
    contribute positively to the field of evolutionary computation. With a focus on modularity and extensibility, it is 
    expected to offer a user-friendly and adaptable environment. It may aid in simplifying evolutionary algorithm 
    implementations, potentially leading to more efficient research and innovation in the field.

    Additionally, we hypothesize that creating a generalized crossover algorithm adaptable to multiple input and output 
    sizes will further enrich the framework's capabilities. Moreover, we believe that integrating advanced and 
    extensible monitoring tools and robust, expressive data validation mechanisms will accelerate evolutionary 
    algorithm development by effectively separating and managing the algorithm's core functions from its monitoring and 
    validation aspects.
    

  \subsection{Research Questions}
    To assess the validity of our hypothesis and further investigate the potential benefits of the proposed
    \textit{Kotlin}-based framework, we outline the following pivotal research questions:

    \begin{enumerate}
      \item What constitute the \textit{fundamental requirements} for a versatile, efficient, and user-friendly
        evolutionary algorithms framework?
      \item How can the features of a language with a \textit{rich type system} --such as static typing, type inference, 
        generics, and null safety--be leveraged to fulfill these requirements and facilitate the development and use of
        evolutionary algorithms?
      \item How does the \textit{ease of use} of the framework compare to that of existing solutions, particularly in
        terms of syntactical complexity?
      \item In what ways can the proposed framework be applied to \textit{address complex real-world problems}, thereby 
        demonstrating its practical viability and potential to contribute to advancements in AI and evolutionary 
        computation?
      \item What potential opportunities exist for the \textit{enhancement and expansion} of the proposed evolutionary 
        algorithm framework in the future?
    \end{enumerate}
    
    These research questions serve as guideposts for our exploration into the feasibility and potential impact of 
    focusing on extensibility and modularity to develop a novel evolutionary algorithms framework. Each question targets
    a specific facet of the project, offering a holistic understanding of the technical intricacies, practical 
    applications, and future prospects of our proposed framework.