\section{Motivation}
\label{sec:motivation}
  The inception of artificial intelligence (AI) as we perceive it today is 
  deeply rooted in the mid-20th century when Alan Turing proposed the intriguing 
  question, \enquote{Can machines think?}~\autocite{turingCOMPUTINGMACHINERYINTELLIGENCE1950a}.
  This seminal query catalyzed a remarkable evolution in the field, leading to the emergence of a diverse spectrum of AI applications that persistently extend and advance.

  Among the early endeavors to emulate evolution artificially, Nills Aall 
  Barricelli's initiative in 1954 stands out.
  Barricelli construed evolution as a purely statistical process, setting a 
  novel precedent for the field of evolutionary computation 
  (EC)~\autocite{barricelliNumericalTestingEvolution1962}.
  This groundbreaking exploration established the parameters for an evolutionary 
  system, which necessitated the system's components to exhibit capabilities of 
  reproduction and mutation, thereby facilitating evolution through natural 
  selection or the survival of the fittest.

  Expanding upon these foundational principles, John Holland's influential book, 
  \emph{Adaptation in Natural and Artificial 
  Systems}~\autocite{hollandAdaptationNaturalArtificial1992a}, unveiled genetic 
  algorithms (GA).
  This innovative methodology for addressing optimization problems drew 
  inspiration from natural selection and genetic mechanisms.
  Holland's work solidified the field of evolutionary algorithms (EA), igniting 
  the development of a plethora of techniques and algorithms, including but not 
  limited to, genetic algorithms, genetic programming, and evolutionary 
  strategies.

  Despite the proven efficacy of evolutionary algorithms in resolving intricate 
  optimization problems ranging from scheduling and data mining to machine 
  learning and beyond, their execution can be challenging, frequently demanding 
  extensive, repetitive coding.
  As the uptake of AI continues to soar, the requirement for streamlined, 
  intuitive libraries and frameworks that empower efficient experimentation and 
  application of these potent algorithms grows concomitantly.

  The \textit{Kotlin} programming language emerges as a promising platform for 
  developing such a framework~\autocite{KotlinProgrammingLanguagea}.
  \textit{Kotlin} distinguishes itself with its lucid syntax, static typing, and 
  flawless interoperability with Java, making it an attractive choice for AI and
  EC enthusiasts~\autocite{CallingJavaKotlin,CallingKotlinJava,KotlinJavaInteropGuide}.
  The significance of \textit{Kotlin} has been formally acknowledged in 
  \textit{Android} app development~\autocite{AndroidKotlinfirstApproach}, and it
  continues to gain popularity among developers, with over 60\% of 
  \textit{Android} developers now using it~\autocite{LatestProgrammingLanguages2023}.
  Furthermore, according to a \textit{GitHub} study, \textit{Kotlin} emerged as 
  the ninth fastest-growing language worldwide from 2021 to 2023, overtaking 
  \textit{Python} by 0.4\%~\autocite{TopProgrammingLanguages}.

  The allure of \textit{Kotlin} is further amplified by its multiplatform 
  feature, which permits code sharing across a wide range of platforms, 
  including JVM, \textit{JavaScript}, \textit{Android}, \textit{iOS}, and native 
  desktop applications.
  This cross-platform compatibility promotes enhanced productivity and 
  consistency in software development.

  \textit{Kotlin} also supports functional programming (FP) and coroutines, 
  which simplify the development of asynchronous and concurrent applications.
  This functionality proves particularly beneficial in the realm of evolutionary 
  algorithms that often necessitate the execution of numerous concurrent 
  processes.
  Moreover, \textit{Kotlin}'s capabilities align well with the design of 
  \textit{Domain Specific Languages} (DSLs)~\autocite{TypesafeBuildersKotlin}, 
  which can be used to concisely and intuitively define the problem domain.

  Given these compelling features, this thesis aims to exploit the prowess of 
  \textit{Kotlin} to architect and implement a pioneering genetic algorithms 
  framework, thereby augmenting the ever-growing application of \textit{Kotlin} 
  within the AI community.
  We expect this framework to stimulate research and applications in the field, 
  hastening the creation of novel algorithms and techniques.
