\section{Objectives}
\label{sec:objectives}
  The primary objectives of this thesis are set forth to establish a solid 
  foundation for the design, implementation, and evaluation of a novel, 
  efficient, and user-friendly genetic algorithms framework using the 
  \textit{Kotlin} programming language.

  \subsection{Main Objective}
    The main objective of this study is to develop a \textit{Kotlin}-based 
    genetic algorithms framework that is versatile, efficient, and
    user-friendly.
    The framework should be able to support classical GAs, as seen in 
    \vref{sec:bg:ga}, and be easily extensible to support future enhancements
    and expansions such as new algorithms, like GP, and genetic operators.

  \subsection{Specific Objectives}
    Building upon the primary objectives, the following specific objectives have
    been established to guide this research:
    
    \begin{enumerate}
      \item \textbf{Language Feature Utilization:} Examine and utilize 
        \textit{Kotlin}'s unique language features, such as its expressive
        syntax, static typing, and its ease of developing DSLs, to design and
        implement a genetic algorithms framework.
        This exploration should enhance the design and implementation of the 
        genetic algorithms framework, ensuring it is efficient, robust, and 
        user-friendly.
      \item \textbf{Framework Efficiency:} Design the genetic algorithms 
        framework to optimize computational efficiency.
        This should involve algorithmic improvements, effective use of 
        \textit{Kotlin}'s language features, and careful resource management.
      \item \textbf{Comparative Study:} Conduct comparative studies between the 
        \textit{Kotlin}-based genetic algorithms framework and existing 
        solutions, particularly in terms of syntactical complexity, and ease of 
        use.
        This should provide a benchmark to evaluate the advantages and potential 
        areas of improvement for our framework.
      \item \textbf{Real-world Applications:} Apply the proposed 
        \textit{Kotlin}-based framework to address complex real-world problems.
        This will not only demonstrate the practical viability of the framework 
        but also its potential to contribute to advancements in AI and EC.
      \item \textbf{Future Enhancements and Expansion:} Anticipate and propose 
        potential enhancements and extensions for the \textit{Kotlin}-based 
        genetic algorithms framework.
        Discuss its adaptability to incorporate future advancements in AI and 
        EC, ensuring its sustainability and continued relevance.
    \end{enumerate} 
    
    These specific objectives aim to add more granularity to our research focus, 
    ensuring a comprehensive and detailed exploration of the possibilities 
    offered by \textit{Kotlin} for the development of a novel genetic algorithms 
    framework.