% Copyright (c) 2023 Ignacio Slater Muñoz All rights reserved.
% Use of this source code is governed by a BSD-style
% license that can be found in the LICENSE file.

\subsubsection{Generation Methods}
\label{sec:keen:gp:primitives:gen}

    Every individual within the population should consistently be a valid program throughout the evolution process. To 
    guarantee the syntactic correctness of 
    all newly created or varied individuals, the GP algorithm must generate an 
    initial population of syntactically accurate individuals. The designated 
    algorithm for producing this initial population is termed the 
    \textit{generation method} or \textit{generator function}.

    Koza's influential work~\autocite{kozaGeneticProgrammingProgramming1992a} 
    introduced two pivotal algorithms for generating the initial population: the 
    \textit{full} and \textit{grow} methods.

    The \emph{full} method creates a tree with a specified maximum depth \(D\), 
    ensuring that all terminal nodes are located at depth \(D\).

    Contrastingly, the \emph{grow} method crafts a tree with a maximum depth of 
    \(D\) and a minimum depth of \(d\), positioning all terminal nodes at depths 
    ranging between \(d\) and \(D\). This method was elaborated upon in 
    \vref{sec:bg:gp:init}.

    It's noteworthy that the \emph{full} method can be viewed as a specific 
    instance of the \emph{grow} method when \(d = D\). Given this perspective, we 
    can further conceptualize the \emph{grow} method as a
    \enquote{depth-conditional tree generation} strategy. The detailed algorithm 
    for this is presented in \vref{lst:keen:gp:primitives:gen}.

    \begin{code}{Depth-conditional tree generation}{
        label=lst:keen:gp:primitives:gen
    }{kotlin}
        fun generateRecursive(
            intermediates: List<Intermediate>,
            leafs: List<Leaf>,
            depth: Int,
            maxHeight: Int,
            condition: (maxHeight: Int, depth: Int) -> Boolean,
            leafFactory: (Leaf) -> Tree,
            intermediateFactory: (Intermediate, List<Tree>) -> Tree,
        ) {
            val children = []
            val node = if (condition(maxHeight, depth)) {
                leafFactory(random node from leafs)
            } else {
                val intermediate = random node from intermediates
                repeat intermediate.arity times {
                    children += generateRecursive(
                        intermediates,
                        leafs,
                        depth + 1,
                        maxHeight,
                        condition,
                        leafFactory,
                        intermediateFactory
                    )
                }
                intermediateFactory(intermediate, children)
            }
            return node
        }
    \end{code}

    The \texttt{generateRecursive} function accepts a list of intermediate nodes, 
    a list of leaf nodes, a maximum height, a current depth, and a condition 
    function. The condition function receives the maximum height and the current 
    depth of a node, returning a boolean value. If the condition is met (true), a 
    leaf node is formed; if not (false), an intermediate node is created. This 
    function is invoked recursively to form the descendants of the current node, 
    halting the recursion when the condition is true or when the maximum height is 
    attained. Additionally, the \texttt{generateRecursive} function takes in two 
    other functions dedicated to the creation of both leaf and intermediate nodes, 
    enabling its adaptability for any tree type, including program trees.

    A key characteristic of this algorithm is its bottom-up approach to tree generation, starting from terminal nodes and 
    finishing at the root node. This method allows for the creation of immutable trees, eliminating the need to make 
    temporary trees.

    The \texttt{generate} function operates by calling the \texttt{generateRecursive} function, with a maximum height randomly determined between \(d\) and \(D\).

    Using this method, the grow and full techniques can be illustrated as:

    \begin{code}{Grow and full methods}{
        label=lst:keen:gp:primitives:gen:growfull
    }{kotlin}
        fun <T> generateProgramGrowing(
            terminals: List<Terminal<T>>,
            functions: List<Fun<T>>,
            min: Int,
            max: Int,
        ) = generateProgram(terminals, functions, min, max) { h, d ->
            d == h || d >= min 
                && random.double() < terminals.size / (terminals.size + functions.size)
        }

        fun <T> generateProgramFull(
            terminals: List<Terminal<T>>, functions: List<Fun<T>>, min: Int, max: Int,
        ) = generateProgram(terminals, functions, min, max) { h, d -> h == d }
        
        fun <T> generateProgram(
            terminals: List<Terminal<T>>,
            functions: List<Fun<T>>,
            min: Int,
            max: Int,
            condition: (Int, Int) -> Boolean,
        ) = Tree.generate(
            nodes = terminals to functions,
            depths = min to max,
            condition = condition,
            leafFactory = ::Program,
            intermediateFactory = ::Program
        )
    \end{code}

    The \texttt{generateProgramGrowing} function creates a program tree using lists of terminal and intermediate nodes, 
    along with minimum and maximum heights. It uses these criteria to decide how to build the tree:

    \begin{itemize}
        \item If the current depth is the same as the maximum height, it stops growing the tree.
        \item If the current depth is at least the minimum height and a random number between 0 and 1 is less than the 
            ratio of terminal to intermediate nodes, it also stops growing.
        \item Otherwise, it continues growing the tree.
    \end{itemize}

    The \texttt{generateProgramFull} function is similar, but it only stops growing the tree when the current depth 
    reaches the maximum height.

    Koza also suggested that population members could be derived using a \textit{ramped half-and-half} technique. Here, the initial population is 
    formed utilizing both the \textit{full} and \textit{grow} strategies, with the 
    method for each individual being selected at random. To implement this, 
    \textit{Keen} offers the \texttt{generateProgramWith} function, receiving a 
    list of generation strategies, a compilation of terminals and functions, and a 
    minimum and maximum height. This function randomly chooses a generation method 
    to produce the program tree. By defining the function this way, integrating 
    new generation techniques into the library becomes seamless.

    In this method, the \textit{ramped half-and-half} technique is just a specific use of the 
    \texttt{generateProgramWith} function. This version limits the list of generation strategies to only the 
    \textit{full} and \textit{grow} methods.
    
    We would like to note that this approach is novel to the \textit{Keen} framework, as other GP libraries usually
    limits the generation method to a single strategy (usually the grow ramped half-and-half method).