\chapter{Glossary}
\label{appendix:glossary}
  In this appendix we present a glossary of terms used throughout this document as a reference for
  the reader.

  \section*{A}
    \begin{definition}[Alteration]
    \label{def:alteration}
      See \vref{def:variation}.
    \end{definition}

    \begin{definition}[Alterer]
    \label{def:alterer}
      See \vref{def:variation_operator}.
    \end{definition}

  \section*{C}
    \begin{definition}[Chromosome]
    \label{def:chromosome}
      Representation of a single column of genetic information of a candidate solution to a given
      optimization problem.

      Formally, a chromosome is a vector \(\textbf{c} = (g_1, g_2, \dots, g_n)\), where \(g_i\) is a
      gene (see \vref{def:gene}).
    \end{definition}

  \section*{E}
    \begin{definition}[Evolutionary computation]
    \label{def:evolutionary_computation}
      Family of algorithms for global optimization inspired by the process of natural selection.

      This typically involves processes mimicking natural selection, mutation, recombination, and 
      survival of the fittest. 
      The solutions to a problem are encoded as a set of \enquote{individuals} in a 
      \enquote{population}. 
      Over multiple generations, these individuals are selected and modified (via genetic operators 
      like crossover and mutation) in order to find better solutions.
    \end{definition}

  \section*{F}
    \begin{definition}[Fitness function]
      % No label
      See \vref{def:fitness_function}.
    \end{definition}
    
  \section*{G}
    \begin{definition}[Gene]
    \label{def:gene}
      Representation of a single component of a candidate solution to a given optimization problem.

      Formally, for a multi-dimensional function \(f\), a gene is an element \(g\) in the domain of
      \(f\).
    \end{definition}

    \begin{definition}[Generation]
    \label{def:generation}
      Number of iterations performed by an evolutionary algorithm.
    \end{definition}

    \begin{definition}[Genotype]
    \label{def:genotype}
      Representation of the full genetic information of a candidate solution to a given optimization
      problem.

      Formally, a genotype is a matrix \(\mathbf{G} = (\textbf{c}_1, \textbf{c}_2, \dots, 
      \textbf{c}_n)\), where \(\textbf{c}_i\) is a chromosome (see \vref{def:chromosome}).
    \end{definition}

  \section*{I}
    \begin{definition}[Individual]
    \label{def:individual}
      A candidate solution to a given optimization problem.

      Formally, an individual is a pair \((\mathbf{G}, \mathbf{f})\), where \(\mathbf{G}\) is the
      genotype (see \vref{def:genotype}) and \(\mathbf{f}\) is the fitness value of the
      individual.
    \end{definition}

  \section*{M}
    \begin{definition}[Metaheuristics]
    \label{def:metaheuristic}
      Problem-independent algorithmic method that yields a sufficiently good solution within 
      reasonable time for an optimization problem, especially for complex problems where an exact 
      solution is not crucial.
    \end{definition}
  
  \section*{P}
    \begin{definition}[Phenotype]
    \label{def:phenotype}
      Same as \vref{def:individual}.
    \end{definition}

    \begin{definition}[Population]
    \label{def:population}
      Set of candidate solutions to a given optimization problem.
    \end{definition}

  \section*{S}
    \begin{definition}[Search space]
    \label{def:search_space}
      Set of all candidate solutions to a given optimization problem.
    \end{definition}
    
    \begin{definition}[Selection]
    \label{def:selection}
      Process of selecting individuals from a population using a selection operator.
    \end{definition}

    \begin{definition}[Selection operator]
    \label{def:selection_operator}
      Operator used to select individuals from a population.

      Formally, a selection operator is a function \(\Sigma : \mathbb{P} \times \mathbb{N} \to
      \mathbb{P}\), where \(\mathbb{P}\) is the set of populations and \(\mathbb{N}\) is the set of
      \(\mathbb{N}\) is the set of natural numbers.
    \end{definition}

    \begin{definition}[Selector]
      See \vref{def:selection_operator}.
    \end{definition}

  \section*{V}
    \begin{definition}[Variation]
    \label{def:variation}
      Process of creating new individuals from existing ones by applying variation operators.
    \end{definition}

    \begin{definition}[Variation operator]
    \label{def:variation_operator}
      Operator used to create new individuals from existing ones.

      Formally, a variation operator is a function \(\varphi : \mathbb{P} \times \mathbb{N} \to
      \mathbb{P}\), where \(\mathbb{P}\) is the set of populations and \(\mathbb{N}\) is the set of
      \(\mathbb{N}\) is the set of natural numbers.
    \end{definition}
